Open source symbolic multibody dynamics software

### 2012-02-20 PetersonGedeHubbard2012.tex
Removed qdot inverses; placed them in mass matrix
Distributed negative sign.
Started splitting up sections

\begin{abstract}
  Abstract here.
\keywords{symbolic dynamics software, nonholonomic systems, control}
\end{abstract}

@@

\section{Derivation of linearization algorithm}
\label{derivations}
The goal of the following calculations is to concretely establish the first
order relationships (about an arbitrary equilibrium point) between an arbitrary
choice of independent coordinates and speeds, and the time-derivatives of all
coordinates and speeds (both independent and dependent).  This enables several
common tasks: (1) stability analysis, (2) control system design, (3) control system design with output feedback based
upon quantities that are dependent (i.e., not state variables), (4)
determination of measurement equations for use in Kalman filter design when
measured quantities are at the acceleration level (i.e., accelerometers).  The
formalism presented here is generic enough to cover most examples of
time-varying constrained multibody systems with arbitrary external inputs and
arbitrary specified quantities.

The complete configuration of the system is assumed to be described by
generalized coordinates $\bm{q}\in\bm{R}^n$ and its velocity completely
described by generalized speeds $\bm{u}\in\bm{R}^o$, with $n \ne o$ in general.
All coordinates and speeds, not only the independent ones, are represented by
$\bm{q}$ and $\bm{u}$.  The system is assumed to be governed by the
relationships described in table \ref{table:assumptions}.  The system is
described by $l + m + n + o$ equations which relate $2n + 2o + 2p + 1$
quantities, which implies that $n - l + o - m + p + 1$ of these quantities are
independent.  This paper discusses the first order relationship between these
independent quantities ($n-l$ independent coordinates, $o-m$ independent
speeds, $p$ control inputs, and time $t$) and the time derivatives of the $n$
coordinates and $o$ speeds.
%It is assumed that there are $l$
%configuration constraints (), and $m$ velocity constraints , and $m$ acceleration
%constraints (). Two functions $ relate the coordinate time derivatives to the generalized speeds,
%and two more functions $\bm{f}_2 : \mathbf{R}^n \times \mathbf{R}^o \times
%\mathbf{R} \to \mathbf{R}^{o-m}$ and $\bm{f}_3 : \mathbf{R}^n
%\times \mathbf{R}^n \times \mathbf{R}^o \times \mathbf{R}^p \times \mathbf{R}
%\to \mathbf{R}^{o-m}$ relate the time derivatives of the speeds $\bm{\dot{u}}$ to
%the externally applied inputs $\bm{r}\in\bm{R}^q$, coordinates $\bm{q}$ and
%their time derivatives $\bm{\dot{q}}$,  speeds $\bm{u}$, and time $t$. These
%assumptions can be summarized by the following equations
%\begin{align}
%  \label{eq:configuration}
%  \bm{f}_{c}(\bm{q}, t) &= \bm{0}\\
%  \label{eq:velocity}
%  \bm{f}_{v}(\bm{q}, \bm{u}, t) &= \bm{0}\\
%  \label{eq:acceleration}
%  \bm{f}_{a}(\bm{q}, \bm{\dot{q}}, \bm{u}, \bm{\dot{u}}, t) &= \bm{0}\\
%  \label{eq:kindiffs}
%  \bm{f}_{0}(\bm{q}, \bm{\dot{q}}, t)
%    + \bm{f}_{1}(\bm{q}, \bm{u}, t) &= \bm{0} \\
%  \label{eq:dyndiffs}
%  \bm{f}_{2}(\bm{q}, \bm{\dot{u}}, t)
%    + \bm{f}_{3}(\bm{q}, \bm{\dot{q}}, \bm{u}, \bm{r}, t) & = \bm{0}
%\end{align}
\begin{table}[htbp]
  \centering
  \caption{Constrained multibody system governing equations}
  \begin{tabular}[c]{r @{ $=$ } l l}
    \multicolumn{2}{c}{Equation} & Mathematical definition \\
    \hline
    $\bm{f}_{c}(\bm{q}, t)$ & $\bm{0}$ & $\bm{f}_c : \mathbf{R}^n \times \mathbf{R} \to
\mathbf{R}^l$ \\
$\bm{f}_{v}(\bm{q}, \bm{u}, t)$ & $\bm{0}$ & $\bm{f}_v : \mathbf{R}^n \times
\mathbf{R}^o \times \mathbf{R} \to \mathbf{R}^m$ \\
$\bm{f}_{a}(\bm{q}, \bm{\dot{q}}, \bm{u}, \bm{\dot{u}}, t)$ & $\bm{0}$ &
$\bm{f}_a : \mathbf{R}^n \times \mathbf{R}^n \times \mathbf{R}^o \times
\mathbf{R}^o \times \mathbf{R} \to \mathbf{R}^m$ \\
$\bm{f}_{0}(\bm{q}, \bm{\dot{q}}, t) + \bm{f}_{1}(\bm{q}, \bm{u}, t)$ &
$\bm{0}$ & $\bm{f}_0 : \mathbf{R}^n \times \mathbf{R}^n \times \mathbf{R} \to
\mathbf{R}^n$\\
\multicolumn{2}{c}{} & $\bm{f}_1 : \mathbf{R}^n \times \mathbf{R}^o \times \mathbf{R}
\to \mathbf{R}^n$ \\
$\bm{f}_{2}(\bm{q}, \bm{\dot{u}}, t) + \bm{f}_{3}(\bm{q}, \bm{\dot{q}}, \bm{u}, \bm{r}, t)$ &
$\bm{0}$ & $\bm{f}_2 : \mathbf{R}^n \times \mathbf{R}^o \times \mathbf{R} \to
\mathbf{R}^{o-m}$,\\
\multicolumn{2}{c}{} & $\bm{f}_3 : \mathbf{R}^n \times \mathbf{R}^n \times
\mathbf{R}^o \times \mathbf{R}^p \times \mathbf{R} \to \mathbf{R}^{o-m}$
  \end{tabular}
  \label{table:assumptions}
\end{table}
The first three equations in table \ref{table:assumptions} represent
constraints derived purely from kinematic considerations at the configuration,
velocity, and acceleration levels, respectively.  The velocity constraints may
be nonholonomic or differentiated holonomic, while the acceleration constraints
are assumed to be differentiated versions of the velocity constraints.  The
fourth equation represents kinematic differential equations which relate
$\bm{\dot{q}}$ and $\bm{u}$ and will be linear in both of these quantities.
The fifth equation represents the constraint free nonholonomic dynamic
differential equations~\cite{Kane1985}, and will necessarily by linear in
$\bm{\dot{u}}$ and input forces and torques $\bm{r}$, but generally nonlinear
in all other quantities.  This formalism is exceedingly generic; in practice,
the structure of the functions in table \ref{table:assumptions} can be
leveraged to simplify the following calculations.  This will be discussed
subsequently.

@@

@@

Equating the right hand sides of equations (\ref{eq:f0_plus_f1_expansion}),
(\ref{eq:acceleration_expansion}),
and (\ref{eq:f2_plus_f3_expansion}) to zero (as per table
\ref{table:assumptions}), and introducing the following definitions
\begin{align}
\tilde{M}_{qq}  &\triangleq \nabla_{\bm{\dot{q}}}\bm{f}_0 \\
\tilde{M}_{uuc} &\triangleq \nabla_{\bm{\dot{u}}}\bm{f}_a \\
\tilde{M}_{uud} &\triangleq \nabla_{\bm{\dot{u}}}\bm{f}_2 \\
\tilde{A}_{qq}  &\triangleq -\nabla_{\bm{q}}(\bm{f}_0 + \bm{f}_1) \\
\tilde{A}_{qu}  &\triangleq -\nabla_{\bm{u}}\bm{f}_1 \\
\tilde{A}_{uqc} &\triangleq -\left[ \nabla_{\bm{\dot{q}}}\bm{f}_a \tilde{M}_{qq}^{-1} \tilde{A}_{qq} + \nabla_{\bm{q}} \bm{f}_a\right] \\
\tilde{A}_{uuc} &\triangleq -\left[ \nabla_{\bm{\dot{q}}}\bm{f}_a \tilde{M}_{qq}^{-1} \tilde{A}_{qu} + \nabla_{\bm{u}} \bm{f}_a\right] \\
\tilde{A}_{uqd} &\triangleq -\left[ \nabla_{\bm{\dot{q}}}\bm{f}_3 \tilde{M}_{qq}^{-1} \tilde{A}_{qq} + \nabla_{\bm{q}} (\bm{f}_2 + \bm{f}_3)\right] \\
\tilde{A}_{uud} &\triangleq -\left[ \nabla_{\bm{\dot{q}}}\bm{f}_3 \tilde{M}_{qq}^{-1} \tilde{A}_{qu} + \nabla_{\bm{u}} \bm{f}_3\right] \\
\tilde{B}_{u}   &\triangleq -\nabla_{\bm{r}}\bm{f}_{3}
\end{align}
enables the unconstrained linear state space equations to be written as
\begin{align}
  \label{eq:state_space_unconstrained}
  \left[
    \begin{array}{cc}
      \tilde{M}_{qq} & \bm{0}_{n \times o} \\
      \bm{0}_{m \times n} & \tilde{M}_{uuc} \\
      \bm{0}_{(o-m) \times n} & \tilde{M}_{uud}
    \end{array}
    \right]
    \left[
      \begin{array}{c}
        \delta \bm{\dot{q}} \\
        \delta \bm{\dot{u}}
      \end{array}
    \right]
   &=
   \left[
     \begin{array}{cc}
       \tilde{A}_{qq} & \tilde{A}_{qu} \\
       \tilde{A}_{uqc} & \tilde{A}_{uuc} \\
       \tilde{A}_{uqd} & \tilde{A}_{uud}
     \end{array}
   \right]
    \left[
      \begin{array}{c}
        \delta \bm{q} \\
        \delta \bm{u}
      \end{array}
    \right]
    +
    \left[
      \begin{array}{c}
        \bm{0}_{(n + m) \times q} \\
        \tilde{B}_{u}
      \end{array}
    \right]
    \delta \bm{r}
\end{align}
Equation (\ref{eq:state_space_unconstrained}) has a state space of dimension $n
+ o$, yet only $p = n - l + o - m$ of these quantities are independent.
To address this issue, a particular set of independent
coordinates and speeds must be selected. To this end, consider the following
partitioning of the generalized coordinates and generalized speeds:
\begin{equation*}
  \tilde{\bm{q}} \triangleq \left[\begin{array}{cc}\bm{q}_{i} &
      \bm{q}_{d}\end{array}\right]^{T} =  P_{q}^{-1} \bm{q}
      \qquad\qquad
  \tilde{\bm{u}} \triangleq \left[\begin{array}{cc}\bm{u}_{i} &
      \bm{u}_{d}\end{array}\right]^{T} =  P_{u}^{-1} \bm{u}
\end{equation*}
where $P_q \in \mathbf{R}^{n \times n}$ and $P_u \in \mathbf{R}^{o \times o}$
are invertible permutation matrices which map an ordering which has the
independent quantities ($\bm{q}_{i}\in\mathbf{R}^{n-l},\,
\bm{u}_{i}\in\mathbf{R}^{o-m})$ first, followed by the dependent quantities
($\bm{q}_{d}\in\mathbf{R}^{l},\, \bm{u}_{d}\in\mathbf{R}^{m}$) to the
original ordering of the coordinates and speeds.  We use the notation $P_{qi}$
and $P_{qd}$ to denote the first $n-l$ and last $l$ columns of $P_q$,
respectively; similarly, $P_{ui}$ is the first $o-m$ columns of $P_{u}$ while
$P_{ud}$ is the last $m$ columns of $P_u$.

@@

By making use of equations (\ref{eq:delta_q}) and (\ref{eq:delta_u}), we can
rewrite equation (\ref{eq:state_space_unconstrained}) as
\begin{align}
  \label{eq:state_space_constrained}
  \left[
    \begin{array}{cc}
      \tilde{M}_{qq} & \bm{0}_{n \times o} \\
      \bm{0}_{m \times n} & \tilde{M}_{uuc} \\
      \bm{0}_{(o-m) \times n} & \tilde{M}_{uud}
    \end{array}
    \right]
    \left[
      \begin{array}{c}
        \delta \bm{\dot{q}} \\
        \delta \bm{\dot{u}}
      \end{array}
    \right]
   &=
   \left[
     \begin{array}{cc}
       (\tilde{A}_{qq} + \tilde{A}_{qu} C_1 ) C_0 & \tilde{A}_{qu} C_2 \\
       (\tilde{A}_{uqc} + \tilde{A}_{uuc} C_1 ) C_0 & \tilde{A}_{uuc} C_2\\
       (\tilde{A}_{uqd} + \tilde{A}_{uud} C_1 ) C_0 & \tilde{A}_{uud} C_2
     \end{array}
   \right]
    \left[
      \begin{array}{c}
        \delta \bm{q}_i \\
        \delta \bm{u}_i
      \end{array}
    \right]
    +
    \left[
      \begin{array}{c}
        \bm{0}_{(n+m) \times q} \\
        \tilde{B}_{u}
      \end{array}
    \right]
    \delta \bm{r}
\end{align}

There are two requirements for this method to work, namely, both
$\nabla_{\bm{q}}\bm{f}_{c} P_{qd}$ and $\nabla_{\bm{u}} \bm{f}_{v} P_{ud}$
must be nonsingular.  Because of the presence of the permutation matrices, this
implies proper selection of dependent coordinates and dependent speeds is
critical.  The permutation matrices effectively select the columns of the
constraint gradient matrices, and may be selected by examination of the
constraint gradient matrix through a SVD or other numerical techniques; for
small systems it may be possible to choose dependent speeds without resorting
to numerical techniques if the system trajectories stay with the configuration
space where these matrices are nonsingular.

For stability analyses, one must examine the stability of the mapping $\delta
\bm{q}_i \to \bm{\dot{q}}_i$.  This can be done by first solving for
$\delta\bm{\dot{x}} = \left[\delta\bm{\dot{q}},\,\delta\bm{\dot{u}}\right]^{T}$, then selecting
the independent rows of these equations ($n-l+o-m$ of them) to form a square
matrix whose eigenvalues govern the stability of the system about the point of
linearization.

>>>>

\begin{abstract}
  Abstract here.
\keywords{symbolic dynamics software, control, nonholonomic systems,
          constrained systems}
\end{abstract}

@@

The goal of the following calculations is to concretely establish the first
order relationships (about an arbitrary equilibrium point) between an arbitrary
choice of independent coordinates and speeds, and the time-derivatives of all
coordinates and speeds (both independent and dependent).  This enables several
common tasks: (1) stability analysis, (2) control system design, (3) control system design with output feedback based
upon quantities that are dependent (i.e., not state variables), (4)
determination of measurement equations for use in Kalman filter design when
measured quantities are at the acceleration level (i.e., accelerometers).  The
formalism presented here is generic enough to cover most examples of
time-varying constrained multibody systems with arbitrary external inputs and
arbitrary specified quantities.

The complete configuration of the system is assumed to be described by
generalized coordinates $\bm{q}\in\bm{R}^n$ and its velocity completely
described by generalized speeds $\bm{u}\in\bm{R}^o$, with $n \ne o$ in general.
All coordinates and speeds, not only the independent ones, are represented by
$\bm{q}$ and $\bm{u}$.  The system is assumed to be governed by the
relationships described in table \ref{table:assumptions}.  The system is
described by $l + m + n + o$ equations which relate $2n + 2o + 2p + 1$
quantities, which implies that $n - l + o - m + p + 1$ of these quantities are
independent.  This paper discusses the first order relationship between these
independent quantities ($n-l$ independent coordinates, $o-m$ independent
speeds, $p$ control inputs, and time $t$) and the time derivatives of the $n$
coordinates and $o$ speeds.
%It is assumed that there are $l$
%configuration constraints (), and $m$ velocity constraints , and $m$ acceleration
%constraints (). Two functions $ relate the coordinate time derivatives to the generalized speeds,
%and two more functions $\bm{f}_2 : \mathbf{R}^n \times \mathbf{R}^o \times
%\mathbf{R} \to \mathbf{R}^{o-m}$ and $\bm{f}_3 : \mathbf{R}^n
%\times \mathbf{R}^n \times \mathbf{R}^o \times \mathbf{R}^p \times \mathbf{R}
%\to \mathbf{R}^{o-m}$ relate the time derivatives of the speeds $\bm{\dot{u}}$ to
%the externally applied inputs $\bm{r}\in\bm{R}^q$, coordinates $\bm{q}$ and
%their time derivatives $\bm{\dot{q}}$,  speeds $\bm{u}$, and time $t$. These
%assumptions can be summarized by the following equations
%\begin{align}
%  \label{eq:configuration}
%  \bm{f}_{c}(\bm{q}, t) &= \bm{0}\\
%  \label{eq:velocity}
%  \bm{f}_{v}(\bm{q}, \bm{u}, t) &= \bm{0}\\
%  \label{eq:acceleration}
%  \bm{f}_{a}(\bm{q}, \bm{\dot{q}}, \bm{u}, \bm{\dot{u}}, t) &= \bm{0}\\
%  \label{eq:kindiffs}
%  \bm{f}_{0}(\bm{q}, \bm{\dot{q}}, t)
%    + \bm{f}_{1}(\bm{q}, \bm{u}, t) &= \bm{0} \\
%  \label{eq:dyndiffs}
%  \bm{f}_{2}(\bm{q}, \bm{\dot{u}}, t)
%    + \bm{f}_{3}(\bm{q}, \bm{\dot{q}}, \bm{u}, \bm{r}, t) & = \bm{0}
%\end{align}
\begin{table}[htbp]
  \centering
  \caption{Constrained multibody system governing definitions and equations}

  \begin{tabular}[c]{l l l}
    Quantity & Shape & Description\\
    \hline
    $\bm{q}$ & $\mathbf{R}^n$ & Coordinates \\
    $\bm{\dot{q}}$ & $\mathbf{R}^n$ & Time derivatives of coordinates \\
    $\bm{u}$ & $\mathbf{R}^o$ & Speeds \\
    $\bm{\dot{u}}$ & $\mathbf{R}^o$ & Time derivatives of speeds \\
    $\bm{r}$ & $\mathbf{R}^p$ & Control inputs \\
  \end{tabular}
  \begin{tabular}[c]{r @{ $=$ } l l}
    \multicolumn{3}{c}{ } \\
    \multicolumn{2}{c}{Equation} & Description \\
    \hline
    $\bm{f}_{c}(\bm{q}, t)$ & $\bm{0}_{l \times 1}$ & Configuration (holonomic)
    constraints \\
    $\bm{f}_{v}(\bm{q}, \bm{u}, t)$ & $\bm{0}_{m \times 1}$ & Velocity
    (nonholonomic) constraints \\
    $\bm{f}_{a}(\bm{q}, \bm{\dot{q}}, \bm{u}, \bm{\dot{u}}, t)$ & $\bm{0}_{m
    \times 1}$ & Acceleration constraints \\
    $\bm{f}_{0}(\bm{q}, \bm{\dot{q}}, t) + \bm{f}_{1}(\bm{q}, \bm{u}, t)$ &
    $\bm{0}_{n \times 1}$ & Kinematic differential equations \\
    $\bm{f}_{2}(\bm{q}, \bm{\dot{u}}, t) + \bm{f}_{3}(\bm{q}, \bm{\dot{q}},
    \bm{u}, \bm{r}, t)$ & $\bm{0}_{(o - m) \times 1}$ & Dynamic differential
    equations \\
  \end{tabular}
  \label{table:assumptions}
\end{table}
The first three equations in table \ref{table:assumptions} represent
constraints derived purely from kinematic considerations at the configuration,
velocity, and acceleration levels, respectively.  The velocity constraints may
be nonholonomic or differentiated holonomic, while the acceleration constraints
are assumed to be differentiated versions of the velocity constraints.  The
fourth equation represents kinematic differential equations which relate
$\bm{\dot{q}}$ and $\bm{u}$ and will be linear in both of these quantities.
The fifth equation represents the constraint free nonholonomic dynamic
differential equations~\cite{Kane1985}, and will necessarily by linear in
$\bm{\dot{u}}$ and input forces and torques $\bm{r}$, but generally nonlinear
in all other quantities.  This formalism is exceedingly generic; in practice,
the structure of the functions in table \ref{table:assumptions} can be
leveraged to simplify the following calculations.  This will be discussed
subsequently.

@@

\section{Derivation of linearization algorithm}
\label{derivations}

@@

Equating the right hand sides of equations (\ref{eq:f0_plus_f1_expansion}),
(\ref{eq:acceleration_expansion}),
and (\ref{eq:f2_plus_f3_expansion}) to zero (as per table
\ref{table:assumptions}), and introducing the following definitions
\begin{align}
\tilde{M}_{qq}  &\triangleq \nabla_{\bm{\dot{q}}}\bm{f}_0 \\
\tilde{M}_{uqc} &\triangleq \nabla_{\bm{\dot{q}}}\bm{f}_a \\
\tilde{M}_{uuc} &\triangleq \nabla_{\bm{\dot{u}}}\bm{f}_a \\
\tilde{M}_{uqd} &\triangleq \nabla_{\bm{\dot{q}}}\bm{f}_2 \\
\tilde{M}_{uud} &\triangleq \nabla_{\bm{\dot{u}}}\bm{f}_2 \\
\tilde{A}_{qq}  &\triangleq -\nabla_{\bm{q}}(\bm{f}_0 + \bm{f}_1) \\
\tilde{A}_{qu}  &\triangleq -\nabla_{\bm{u}}\bm{f}_1 \\
\tilde{A}_{uqc} &\triangleq -\tilde{A}_{qq} - \nabla_{\bm{q}} \bm{f}_a \\
\tilde{A}_{uuc} &\triangleq -\tilde{A}_{qu} - \nabla_{\bm{u}} \bm{f}_a \\
\tilde{A}_{uqd} &\triangleq -\tilde{A}_{qq} - \nabla_{\bm{q}} (\bm{f}_2 + \bm{f}_3) \\
\tilde{A}_{uud} &\triangleq -\tilde{A}_{qu} - \nabla_{\bm{u}} \bm{f}_3 \\
\tilde{B}_{u}   &\triangleq -\nabla_{\bm{r}}\bm{f}_{3}
\end{align}
enables the unconstrained linear state space equations to be written as
\begin{align}
  \label{eq:state_space_unconstrained}
  \left[
    \begin{array}{cc}
      \tilde{M}_{qq} & \bm{0}_{n \times o} \\
      \tilde{M}_{uqc} & \tilde{M}_{uuc} \\
      \tilde{M}_{uqd} & \tilde{M}_{uud}
    \end{array}
    \right]
    \left[
      \begin{array}{c}
        \delta \bm{\dot{q}} \\
        \delta \bm{\dot{u}}
      \end{array}
    \right]
   &=
   \left[
     \begin{array}{cc}
       \tilde{A}_{qq} & \tilde{A}_{qu} \\
       \tilde{A}_{uqc} & \tilde{A}_{uuc} \\
       \tilde{A}_{uqd} & \tilde{A}_{uud}
     \end{array}
   \right]
    \left[
      \begin{array}{c}
        \delta \bm{q} \\
        \delta \bm{u}
      \end{array}
    \right]
    +
    \left[
      \begin{array}{c}
        \bm{0}_{(n + m) \times q} \\
        \tilde{B}_{u}
      \end{array}
    \right]
    \delta \bm{r}
\end{align}
Equation (\ref{eq:state_space_unconstrained}) has a state space of dimension $n
+ o$, yet only $p = n - l + o - m$ of these quantities are independent.
To address this issue, a particular set of independent
coordinates and speeds must be selected. To this end, consider the following
partitioning of the generalized coordinates and generalized speeds:
\begin{equation*}
  \tilde{\bm{q}} \triangleq \left[\begin{array}{cc}\bm{q}_{i} &
      \bm{q}_{d}\end{array}\right]^{T} =  P_{q}^{-1} \bm{q}
      \qquad\qquad
  \tilde{\bm{u}} \triangleq \left[\begin{array}{cc}\bm{u}_{i} &
      \bm{u}_{d}\end{array}\right]^{T} =  P_{u}^{-1} \bm{u}
\end{equation*}
where $P_q \in \mathbf{R}^{n \times n}$ and $P_u \in \mathbf{R}^{o \times o}$
are invertible permutation matrices which map an ordering which has the
independent quantities ($\bm{q}_{i}\in\mathbf{R}^{n-l},\,
\bm{u}_{i}\in\mathbf{R}^{o-m})$ first, followed by the dependent quantities
($\bm{q}_{d}\in\mathbf{R}^{l},\, \bm{u}_{d}\in\mathbf{R}^{m}$) to the
original ordering of the coordinates and speeds.  We use the notation $P_{qi}$
and $P_{qd}$ to denote the first $n-l$ and last $l$ columns of $P_q$,
respectively; similarly, $P_{ui}$ is the first $o-m$ columns of $P_{u}$ while
$P_{ud}$ is the last $m$ columns of $P_u$.

@@

By making use of equations (\ref{eq:delta_q}) and (\ref{eq:delta_u}), we can
rewrite equation (\ref{eq:state_space_unconstrained}) as
\begin{align}
  \label{eq:state_space_constrained}
  \left[
    \begin{array}{cc}
      \tilde{M}_{qq} & \bm{0}_{n \times o} \\
      \tilde{M}_{uqc} & \tilde{M}_{uuc} \\
      \tilde{M}_{uqd} & \tilde{M}_{uud}
    \end{array}
    \right]
    \left[
      \begin{array}{c}
        \delta \bm{\dot{q}} \\
        \delta \bm{\dot{u}}
      \end{array}
    \right]
   &=
   \left[
     \begin{array}{cc}
       (\tilde{A}_{qq} + \tilde{A}_{qu} C_1 ) C_0 & \tilde{A}_{qu} C_2 \\
       (\tilde{A}_{uqc} + \tilde{A}_{uuc} C_1 ) C_0 & \tilde{A}_{uuc} C_2\\
       (\tilde{A}_{uqd} + \tilde{A}_{uud} C_1 ) C_0 & \tilde{A}_{uud} C_2
     \end{array}
   \right]
    \left[
      \begin{array}{c}
        \delta \bm{q}_i \\
        \delta \bm{u}_i
      \end{array}
    \right]
    +
    \left[
      \begin{array}{c}
        \bm{0}_{(n+m) \times q} \\
        \tilde{B}_{u}
      \end{array}
    \right]
    \delta \bm{r}
\end{align}

\section{Implementation and Examples}

There are a number of additional steps needed to use these linear equations or
form them more efficiently. There are also non-singularity requirements
involved in selection of independent quantities.

In order to use these equations in the state space representation, 2 things
need to happen. First, matrices $\tilde{A}$ and $\tilde{B}$ need to be
multiplied by the mass matrix inverse. We can then define

\begin{align}
    \prime{A} &\triangleq \tilde{M}^{-1} \tilde{A} \\
    \prime{B} &\triangleq \tilde{M}^{-1} \tilde{B}
\end{align}

Now we will have matrices $\prime{A}$ and $\prime{B}$ which are of size $(o +
n) \times (o - m + n - l)$ and $(o + n) \times p)$. We need to now extract only
the rows which are associated with the independent quantities. We will define
% Is there a better way to write transforming A' to A, B' to B?
% Could we use the permutation matrices here?
the reduced matrices as $A$ and $B$. We will also define the state space vector
$x = \left[\delta\bm{q}_i,\,\delta\bm{u}_i\right]^{T}$; we now have the
traditional state space representation.

%For stability analyses, one must examine the stability of the mapping $\delta
%\bm{q}_i \to \bm{\dot{q}}_i$.  This can be done by first solving for
%$\delta\bm{\dot{x}} = \left[\delta\bm{\dot{q}},\,\delta\bm{\dot{u}}\right]^{T}$, then selecting
%the independent rows of these equations ($n-l+o-m$ of them) to form a square
%matrix whose eigenvalues govern the stability of the system about the point of
%linearization.

Additionally, there are two requirements for this method to work, namely, both
$\nabla_{\bm{q}}\bm{f}_{c} P_{qd}$ and $\nabla_{\bm{u}} \bm{f}_{v} P_{ud}$
must be nonsingular.  Because of the presence of the permutation matrices, this
implies proper selection of dependent coordinates and dependent speeds is
critical.  The permutation matrices effectively select the columns of the
constraint gradient matrices, and may be selected by examination of the
constraint gradient matrix through a SVD or other numerical techniques; for
small systems it may be possible to choose dependent speeds without resorting
to numerical techniques if the system trajectories stay with the configuration
space where these matrices are nonsingular.

The computational cost of computing the jacobian matrices can also be reduced
if the point of linearization is known ahead of time.
